<!-- 
This is a child document for presentation and analysis of RRA results (Mageck)
to be used with screen_report.v0.x.R

 -->



```{r include=FALSE,cache=FALSE}

library(knitr)
library(bookdown)


knitr::opts_chunk$set(echo = FALSE, 
                      collapse = TRUE, 
                      warning = FALSE, 
                      message = FALSE)


source("./shRNA_screen_report_functions.R")

```


### Comparison: `r i` {.tabset}

```{r include=FALSE,cache=FALSE}
indir.i=file.path(mageck.dir,paste(proj.name.pref,i,sep="."),setup)

infile.pth.gene.i=file.path(indir.i,paste(proj.name.pref,i,setup,"gene_summary.txt",sep="."))
infile.pth.sgRNA.i=file.path(indir.i,paste(proj.name.pref,i,setup,"sgrna_summary.txt",sep="."))

gdata_i_lfc = ReadRRA(infile.pth.gene.i,score="lfc")
gdata_i_RRA = ReadRRA(infile.pth.gene.i,score="RRA")


ref.smpl=contrasts.tab$reference[contrasts.tab$name==i]
treat.smpl=contrasts.tab$treatment[contrasts.tab$name==i]


resdir.i=file.path(resdir,i)
dir.create(resdir.i)
plotdir.i=file.path(resdir.i,"plots")
dir.create(plotdir.i)

all.res.rra[[i]]=gdata_i_RRA
```

This section refers to the comparison **`r i` : `r treat.smpl` vs. `r ref.smpl`**


<br />
<br />

```{r include=FALSE, cache=FALSE}
rep.g.i=read.table(infile.pth.gene.i, sep="\t", header=TRUE)
rep.sg.i=read.table(infile.pth.sgRNA.i, sep="\t", header=TRUE)

all.res[[i]]=rep.g.i


```


```{r include=FALSE, cache=TRUE}


# from Hs db
gene_names=data.frame(mapIds(org.Hs.eg.db, keys=rep.g.i$id, keytype="REFSEQ", column="GENENAME", multiVals="first"))
gene_names$id=rownames(gene_names)
colnames(gene_names)=c("description","id")
gene_symbols=data.frame(mapIds(org.Hs.eg.db, keys=rep.g.i$id, keytype="REFSEQ", column="SYMBOL", multiVals="first"))
gene_symbols$id=rownames(gene_symbols)
colnames(gene_symbols)=c("symbol","id")
gene_names=left_join(gene_names,gene_symbols)

gene_entrez=data.frame(mapIds(org.Hs.eg.db, keys=rep.g.i$id, keytype="REFSEQ", column="ENTREZID", multiVals="first"))
gene_entrez$id=rownames(gene_entrez)
colnames(gene_entrez)=c("entrezgene_id","id")

# from library_desc
geneannot=left_join(gene_names,gene_entrez)
shannot=left_join(geneannot,library_desc[,-9], by="id")

```

```{r include=FALSE, cache=FALSE}
rep.g.i.annot=left_join(rep.g.i,geneannot,by="id")


rep.g_neg=rep.g.i.annot[,c(1,16,15,2:8,17)]
rep.g_pos=rep.g.i.annot[,c(1,16,15,2,9:14,17)]

rep.g_neg=rep.g_neg[order(rep.g_neg$neg.rank, decreasing=FALSE),]
rep.g_pos=rep.g_pos[order(rep.g_pos$pos.rank, decreasing=FALSE),]

tab_n_sv=tab_n_sv+1
fname=paste("Table",tab_n_sv,"RRA_mageck",i,setup,"tsv",sep=".")
save.table(df=rep.g.i.annot, file=fname, dir=resdir.i)
```


#### Summary

Results of robust rank aggregation (RRA) analysis (MAGeCK) to identify significantly enriched or depleted sgRNAs are presented separately in sections below.

There are **n= `r nrow(rep.g_neg[rep.g_neg$neg.fdr<=0.05,])` ** hits with **FDR <= 0.05** in the test for **negative selection** in `r i`.

There are **n= `r nrow(rep.g_pos[rep.g_pos$pos.fdr<=0.05,])` ** hits with **FDR <= 0.05** in the test for **positive selection** in `r i`.

N = `r nrow(rep.g.i.annot)` genes were detected in the data.

The entire table can be explored in file **`r fname`**.


<br />
<br />

##### Negative Selection

```{r cache=FALSE}
tab_n=tab_n+1
tablabel=paste(tab_n, "Table", sep="")
```

Results of robust rank aggregation (RRA) analysis (MAGeCK) to identify **depleted genes** are presented in Table \@ref(tab:`r tablabel`). Top 100 rows (by `neg.rank`) are included. 


Fields in the table:

* `id` gene id

* `description` of gene product

* `num` number of shRNA probes targeting the gene

* `score` RRA score by MAGeCK

* `p value` uncorrected p value from the RRA test

* `FDR` false discovery rate, i.e. p value adjusted for multiple comparisons

* `rank` position on a gene list ranked by the RRA score

* `good-shRNA` (originally `goodsgRNA`) number of shRNAs whose ranking is below the alpha cutoff (0 in this case); i.e. number of probes following the expected log2FC for negative selection

* `log2FC` (`lfc` in the source file) gene level log2(fold change)


Fields specific to **negative** selection are prefixed with `neg` in the source table `r infile.pth.gene.i`.


<br />

```{r cache=FALSE, results='asis'}
capt=paste("Summary of gene level analysis of ",i," negative selection (MAGeCK). Shown are top 100 genes (by \"neg.rank\").", sep="")
test_table(capt, label=tablabel)

tbl_report=rep.g_neg[,1:10]
colnames(tbl_report)=c("Id","Symbol","Description","Number of shRNAs per gene","Score","p value","FDR","Rank","good-shRNA","log2FC")
DT::datatable(tbl_report %>% filter(Rank <=100), rownames = FALSE) %>% formatSignif(columns=c('FDR','Score','p value','log2FC'), digits=4)
```





<br />
<br />

##### Positive Selection

```{r cache=FALSE}
tab_n=tab_n+1
tablabel=paste(tab_n, "Table", sep="")
```

Results of robust rank aggregation (RRA) analysis (MAGeCK) to identify **enriched genes** are presented in Table \@ref(tab:`r tablabel`). Top 100 rows (by `Rank`) are included. 

Fields in the table:

* `id` gene id

* `description` of gene product

* `num` number of shRNA probes targeting the gene

* `score` RRA score by MAGeCK

* `p value` uncorrected p value from the RRA test

* `FDR` false discovery rate, i.e. p value adjusted for multiple comparisons

* `rank` position on a gene list ranked by the RRA score

* `good-shRNA` (originally `goodsgRNA`) number of shRNAs whose ranking is below the alpha cutoff (0 in this case);  i.e. number of probes following the expected log2FC for positive selection

* `log2FC` (`lfc` in the source file) gene level log2(fold change)


Fields specific to **positive** selection are prefixed with `pos` in the source table `r infile.pth.gene.i`.



<br />

```{r cache=FALSE, results='asis'}
capt=paste("Summary of gene level analysis of ",i," , positive selection (MAGeCK). Shown are top 100 genes (by \"Rank\").", sep="")
test_table(capt, label=tablabel)
tbl_report=rep.g_pos[,1:10]
colnames(tbl_report)=c("Id","Symbol","Description","Number of shRNAs per gene","Score","p value","FDR","Rank","good-shRNA","log2FC")
DT::datatable(tbl_report %>% filter(Rank <=100), rownames = FALSE) %>% formatSignif(columns=c('FDR','Score','p value','log2FC'), digits=4)
```



#### Interactive Volcano Plot



```{r include=FALSE, warning=FALSE}

dat=gdata_i_lfc
xaxis="Effect size (log2-fold change)"


dat$group="ns"
dat$group[dat$FDR<=FDR.CO] <-"sig"
dat$logFDR=-(log10(dat$FDR))

dat$group[dat$group=="sig" & dat$Score>0] <-"up"
dat$group[dat$group=="sig" & dat$Score<0] <-"dn"

#dat_plot=dat[dat$FDR<=0.8,]

dat_plot=dat

p_v1=ggplot(dat_plot, aes(x=Score, y=logFDR, color=group, text=paste(id,"; Score",round(Score, digits=3),"; FDR",round(FDR, digits=4)))) + 
  geom_point(alpha=0.5) + scale_color_manual(values=mycolour) +
  theme_bw() + theme(legend.position = "none")+
  ylab("-log10(FDR)") + xlab(xaxis)


# geom_vline(xintercept=1.5, linetype="dashed", color = "grey50") + geom_vline(xintercept=-1.5, linetype="dashed", color = "grey50") +  
# geom_hline(yintercept=-log10(0.05), linetype="dashed", color = "grey50")

figurecap=paste("Volcano plot for comparison ",i,". Genes with changes at FDR below the significance cutoff FDR=",FDR.CO, " are coloured. Only genes with FDR<=0.8 are shown.", sep="")


fig_n=fig_n+1
fname=paste("Figure",fig_n,"Volcano",i,setup,"pdf",sep=".")
pdf(file.path(plotdir.i,fname))
p_v1 + ggtitle(figurecap)
dev.off()



p_v1=ggplotly(p_v1, tooltip=c("text"))


```



Volcano plot shows statistical significance as False Discovery Rate, FDR (here shown as -log10(FDR)) versus magnitude of change 
(log2(fold change)).
It enables quick visual identification of genes with large fold changes that are also statistically significant.

The results of abundance analysis at the gene level are visualised on the interactive volcano plot where the statistically significant genes
(FDR<=0.05) are coloured. 


This plot is saved as a static plot in file

 `r paste0(fname, " .")`



```{r dev='png', fig.show='show',fig.dim=c(6, 6), cache=FALSE, fig.cap=figurecap}
p_v1
```


#### Rank Plots

##### Gene Level

```{r include=FALSE}

dat=gdata_i_lfc



dat$group="ns"
dat$group[dat$FDR<=FDR.CO] <-"sig"
dat$logFDR=-(log10(dat$FDR))

dat$group[dat$group=="sig" & dat$Score>0] <-"up"
dat$group[dat$group=="sig" & dat$Score<0] <-"dn"



dat$Rank = rank(dat$Score)

#dat_plot=dat[dat$FDR<=0.8,]
dat_plot=dat

p4_v1=ggplot(dat_plot, aes(x=Score, y=Rank, color=group, text=paste(id,"; Score",round(Score, digits=3),"; Rank",Rank))) +  geom_point(alpha=0.5) + scale_color_manual(values=mycolour) +
theme_bw() + theme(legend.position = "none")# +
# geom_vline(xintercept=1.5, linetype="dashed", color = "grey50") + geom_vline(xintercept=-1.5, linetype="dashed", color = "grey50") +  
# geom_hline(yintercept=-log10(0.05), linetype="dashed", color = "grey50")


figurecap=paste("Gene rank plot for comparison ",i,". Genes with changes at FDR below the significance cutoff FDR=",FDR.CO, " are coloured. Only genes with FDR<=0.8 are shown.", sep="")


fig_n=fig_n+1
fname=paste("Figure",fig_n,"GeneRank",i,setup,"pdf",sep=".")
pdf(file.path(plotdir.i,fname))
p4_v1+ ggtitle(figurecap)
dev.off()


p4_v1=ggplotly(p4_v1, tooltip=c("text"))

```


Gene rank plot depicts the relationship of gene rank and its fold change in treatment vs control (presented as log2(fold change), log2FC).


This plot is interactive. Its static version is saved in file `r fname`.


```{r dev='png', echo=FALSE, fig.show='show',fig.dim=c(6, 6), cache=FALSE, fig.cap=figurecap}
p4_v1
```



<br />
<br />

##### shRNA Level



```{r include=FALSE, warning=FALSE}

sdata_rra = ReadsgRRA(infile.pth.sgRNA.i)
p2 = sgRankView(sdata_rra, top = 5, bottom = 5)


sdata_symbol=sdata_rra
colnames(sdata_symbol)=c("sgrna","id","LFC","FDR")
sdata_symbol=left_join(sdata_symbol,gene_symbols)
sdata_symbol$Gene=paste(sdata_symbol$symbol,sdata_symbol$id,sep="..")
sdata_symbol=sdata_symbol[,c(1,6,3,4)]
colnames(sdata_symbol)=c("sgrna","Gene","LFC","FDR")


#sdata=sdata_rra
sdata=sdata_symbol

figurecap1=paste("shRNA ranks for comparison ",i,".\nTop genes in negative selection (by neg.rank) are shown (top genes at the bottom of the plot).\n shRNAs are coloured by their fold change direction.", sep="")
figurecap2=paste("shRNA ranks for comparison ",i,".\nTop genes in positive selection (by pos.rank) are shown (top genes at the bottom of the plot).\n shRNAs are coloured by their fold change direction.", sep="")

  # here do this for top 10 and save to files for top 20
  sel.dat=rep.g_neg %>% arrange(neg.rank) %>% filter(neg.rank <=20)
  sel.genes_join=paste(sel.dat$symbol[1:10],sel.dat$id[1:10],sep="..")
  p2.1 = sgRankView(sdata, gene=sel.genes_join, top=0, bottom=0)

  sel.genes_join=paste(sel.dat$symbol,sel.dat$id,sep="..")
  p2.3 = sgRankView(sdata, gene=sel.genes_join, top=0, bottom=0) + ggtitle(figurecap1)


  sel.dat=rep.g_pos %>% arrange(pos.rank) %>% filter(pos.rank <=20)
  sel.genes_join=paste(sel.dat$symbol[1:10],sel.dat$id[1:10],sep="..")
  p2.2 = sgRankView(sdata, gene=sel.genes_join, top=0, bottom=0)
  sel.genes_join=paste(sel.dat$symbol,sel.dat$id,sep="..")
  p2.4 = sgRankView(sdata, gene=sel.genes_join, top=0, bottom=0) + ggtitle(figurecap2)



  graphics.off()
  fig_n=fig_n+1
  ggsave(filename=paste("Figure",fig_n,"shRNA_ranks_positive",i,setup,"pdf",sep="."),plot=p2.4,path=plotdir.i,device="pdf")

  graphics.off()
  fig_n=fig_n+1
  ggsave(filename=paste("Figure",fig_n,"shRNA_ranks_negative",i,setup,"pdf",sep="."),plot=p2.3,path=plotdir.i,device="pdf")


```



Visualisation of ranks of individual shRNAs targeting each gene adds more information to gene level results.

The guide rank plots visualize the rank of shRNAs targeting top genes (n=10) under negative and positive selection. 
Rank plots for top 20 genes changed in each direction are saved in file
 `r paste0(fname, " .")`



```{r dev='png', echo=FALSE, fig.show='show',fig.dim=c(6, 6), cache=FALSE, fig.cap=figurecap1}
p2.1
```


```{r dev='png', echo=FALSE, fig.show='show',fig.dim=c(6, 6), cache=FALSE, fig.cap=figurecap2}
p2.2
```





